---
title: "Aides Drug"
author: "Adam Ayala"
date: "Fall 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
female<-filter(triUmp, gender=="F")
```

```{r}
barchartGC(~infected + treatment,data=triUmp)
barchartGC(~infected + treatment,data=triUmp, type="percent")

```

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")

```

The visual data shows that the drug worked when it includes all that took the drug compared to all who took the placebo. But once we see the females only, we can see that the drug actually increased aids and the placebo had a lower percentage of those who got aids.



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=female)
rowPerc(table1)
colPerc(table1)
```

The percentage in the group that didn't get aids is close to 41%. When you look at the people who got aids it is shown to be near 59%, a 18% difference.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value is 0.00007686 for the first Chi square test. Then the next Chi square test p-value is 0.0001. Finally the Fishers Exact test has a p-value of 0.00006596. These values show that there is statistical difference from the data and proves something is happening.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The visual data showed something might be going on for females since they had a higher percentage of the drug giving aids compared to the placebo while the whole amount of people being recorded showed a lower percentage than the placebo. Using numerical data we could see a 11% difference from the female and the original data of how much more got aids from the drug. Then our inferential results proved that something was happening between the 2 since the p-value of the first Chi square test is 0.00007686, then the next Chi square test p-value is 0.0001 and Fishers Exact test has a p-value of 0.00006596. Showing that something is happening and that something is the difference between the female group and the rest of the group when it comes to the effects of taking the drug as it shows to actual give aids when the drug is taken with a 11% higher rate of infection.
